---
title: "Assignment 4: Exploratory Data Analysis"
author: "Greta Bornmann"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Overview

Due: **Friday 9/25 11:59pm**

In this assignment, you will practice Exploratory Data Analysis (EDA) by making plots from a **Kaggle dataset of your choice**.

You will submit **at least 4 different plot types**, with short written interpretations for each.

Part 1 is worth a total of 1 point, and Part 2 is worth 4 points (1 point for each plot, question, and interpretation).

## Part 1: Get Your Data

1.  Go to [Kaggle Datasets](https://www.kaggle.com/datasets), create an account, and choose a dataset that interests you. Ensure your data has:
    1.  At least 4 variables
    2.  At least one categorical and one continuous variable
    3.  I recommend data that contains at least 50-100 observations
    4.  Tip: Open the CSV on Kaggle first and skim the columns to check that it meets these requirements
2.  Download the dataset as a CSV file.
3.  Place the CSV in your class R project `data` folder.
4.  Read it into R with `read_csv()`.

```{r, message=FALSE}
library(tidyverse)

crab_data <- read_csv("data/CrabAgePrediction.csv")

```

### Describe your data

Data author: Gursewak Singh Sidhu

Link: <https://doi.org/10.34740/kaggle/dsv/2834512>

Description: This data set uses variables such as sex, length, diameter, and various weights to predict the age of the crab. This is helpful to understand what features develop with age and decide what is the prime age to harvest crabs.

### Preview your data

Use some of the strategies you've learned in this class or picked up elsewhere to summarize and quickly explore your data.

```{r}

summary(crab_data)
head(crab_data)
tail(crab_data,5)
```

### Summary stats

Before you make plots, you must also:

- Show basic summaries of your dataset, including:

  - `count()` for categorical variables (to see category frequencies)

  - `mean()`, `median()`, `sd()`, `min()`, `max()` for at least one numeric variable

  - Remember to use `na.rm = TRUE` if you have missing values that affect summaries!

```{r}
count(crab_data, Sex)

mean(crab_data$Weight, na.rm = TRUE)
median(crab_data$Weight, na.rm = TRUE)
sd(crab_data$Weight, na.rm = TRUE)
min(crab_data$Weight, na.rm = TRUE)
max(crab_data$Weight, na.rm = TRUE)
```

Write 2–3 sentences here: What do the category counts tell you? What does the numeric summary show about spread, center, or outliers?

The category count told me a large percentage of the crabs' sex was indeterminate, but the female to male ratio was surprisingly equal. The numerica summary showed that the mean of the crab's weight was on the lower side, and the max of 80.1 ounces is a very high outlier compared to the mean of 23.5 ounces. The median being low (22.8 ounces) as well told me that the spread would be on the lower 40% of the data.

## Part 2: Visualize Your Data

Create **at least 4 plots**, each of a different type. Your plots may share variables, but each one should answer a **different question** about your data. At least one plot must include a variable that isn't in the others.

Each plot should have:

- A title and labeled axes

- A short interpretation (2-3 sentences) underneath

  - Sample interpretation: (boxplot of `price` by `cut` from `diamonds` data): *Ideal diamonds have the lowest median price, which is surprising since Ideal is the best cut. Every cut also has a lot of very expensive outliers. One possible reason for the low Ideal prices is that Ideal diamonds might be smaller on average, which I could check by plotting carat by cut.*

Use the EDA Part 3 notes as your reference. Since this is an EDA assignment, your plots can look plain, but consider adding colors, themes, or interactivity for interpretability and visual appeal if you know how.

### Plot 1

```{r}

ggplot(crab_data, aes(x = Sex, y = Weight, fill = "coral")) +
  geom_boxplot() +
  labs(title = "Crab Weight by Sex", x = "sex", y = "weight")

```

Question this plot answers: How does the weight of crabs vary by sex?

Interpretation: It seems that female crabs tend to weigh slightly more on average than the males do, with outlier exceptions. It also seems that the indeterminate crabs tend to be lighter, which makes me think that the smaller (probably younger) crabs make it harder to differentiate sex.

### Plot 2

```{r}
ggplot(crab_data, aes(x = Sex)) +
  geom_bar(fill = "aquamarine") +
  labs(title = "Number of Crabs by Sex", x = "sex", y = "number of crabs")
```

Question this plot answers: How many female, male, and indeterminate crabs were selected.

Interpretation: The majority of the determinable crabs were male, by a couple hundred. There were more crabs that they weren't able to determine the sex of than female crabs which is surprising because its not that hard to tell if a crab is male or female.

### Plot 3

```{r}
ggplot(crab_data, aes(x = Weight, fill = Sex)) +
  geom_histogram() +
  labs(title = "Crab Weight and Age", x = "Weight (ounces)", y = "Age (months)")+
  stat_bin(bins = 50) +
  scale_fill_manual(values = c("F" = "pink","M" = "navy", "I" = "coral"))
```

Question this plot answers: How does crab weight correlate with age and sex?

Interpretation: Male crabs tend to mostly weigh 20 to 40 ounces while females have a much larger ranger. I noticed the indeterminable crabs were smaller but all ages.

### Plot 4

```{r}
ggplot(crab_data, aes(x = Length, y = Weight, color = Sex,)) +
  geom_point() +
  scale_color_manual(
    values = c("F" = "pink","M" = "navy", "I" = "coral"))
  labs(title = "Crabs Weigh vs. Length", x = "length", y = "weight") +
  theme_bw()
```

Question this plot answers: Is crab length related to crab weight?

Interpretation: The male crabs are both longer and weigh more than the other two catagories. The female crabs have less of a pattern and are more scattered along the line of best fit. The indeterminable crabs are on the smaller side for both catagories.

------------------------------------------------------------------------

# Push to GitHub AND submit link on Canvas!!

Assignment is due Friday 9/25 at 11:59 pm.
